\documentclass[tcc]{subfile}

\begin{document}

\begin{agradecimentos}

    I would like to thank my supervisor Prof.\ José Eduardo Mautone de Barros for
    the guidance and support throughout this work, even more so because I was
    flying back and forth between Belo Horizonte and  São José dos Campos 
    and was not able to meet him as often as I would like.

    I would also like to thanks Prof. Luiz Ricardo Utsch de Freitas Pinto, who was
    always had his door opened for a friendly exchange of ideas and has always provided
    me with very enriching discussions, not only during the execution this work
    but also throughout my entire undergraduate course.  In the same spirit, I
    would like to thank Prof.\ Eduardo Bauzer de Medeiros, who always offered
    me a helping hand when I needed one.

    I am also grateful to the people at EMBRAER, both in the Flight Operations
    team and in the Preliminary Design team, for going out of their
    way to ensure that I could conciliate the internship there with my studies
    during these rather short one and a half years. I
    hope I have the opportunity to work with them again.

    I would like to express my appreciation to Prof. Christie Maddock, from the
    University of Strathclyde, for giving me the opportunity of doing research overseas for the first time,
    and also for sparking my interest in the amazing subject of Dynamics and Control.

    Thanks to the people at CEA, whose camaraderie was invaluable during these years. 
    Thanks especialy to Daniel Andrade, who kindly made the technical drawings 
    for the annex of this work.

    Last but not least, I would like to thank my family, who not only had to put up
    with me working for weekends straight to complete this thesis but was also
    pivotal in encouraging and supporting me to become who I am today, both
    academically and personally. 

\end{agradecimentos}

\end{document}
